\vspace{-2.5em} Spatio-angular transfer functions for polarized fluorescence microscopes\vspace{-1em}

### 2018-05-03 notes/2018-04-19-polarizer-otf/report/report.tex
Redefine illumination constants to simplify. Fix signs on ill y_0^2.

We calculated several spatio-angular excitation point response functions in our
Optics Express paper \cite{chandler17} (in the paper we called this function the
excitation efficiency for a single dipole, but this is equivalent to the
excitation point spread function). If we place a spatially incoherent and
spatially uniform light source (or its image) in the aperture plane with the
optical axis along the $\mh{z}$ axis, then the excitation point response function
is given by
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\Theta, \Phi; \phi_{\text{exc}}) &= \bar{D}\{\bar{A} + \bar{B}\sin^{2}{\Theta} + \bar{C}\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs},
\end{align}
where
$\so{} = \cos\Phi\sin\Theta\mh{x} + \sin\Phi\sin\Theta\mh{y} + \cos\Theta\mh{z}$,
the illumination polarizer orientation is given by $\mh{p} = \cos\phi_{\text{exc}}\mh{x} + \sin\phi_{\text{exc}}\mh{y}$, and 
\begin{subequations}
\begin{align}
  \bar{A} &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha },\\
  \bar{B} &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha },\\
  \bar{C} &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha},\\
  \bar{D} &= \frac{4}{3(1 - \cos\alpha)},
\end{align}\label{eq:coefficients}%
\end{subequations}
where $\alpha = \arcsin(\text{NA}/n_o)$. I'm using bars on the constants to
avoid notation overlap with the previous note set. We can rewrite this
expression in terms of spherical harmonics as
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p}) \propto\, &[2\bar{A} + (4/3)\bar{B}]\bar{D}y_0^0(\so{}) - \frac{4\sqrt{5}}{15}\bar{B}\bar{D}y_2^0(\so{}) + \nonumber\\ &\frac{4\bar{C}\bar{D}}{\sqrt{15}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}. \label{eq:genpsf}
\end{align}

The light-sheet excitation point response function is given by Eq.
\ref{eq:genpsf} in the limit of small NA which gives
\begin{align}
  h^{\mh{z}, \text{ls}}_{\text{exc}}(\so{}; \mh{p}) \equiv \lim_{\alpha \rightarrow 0} h^{\mh{z}}_{\text{exc}}(\so{}) \propto y_0^0(\so{}) + \frac{1}{\sqrt{5}}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}.
\end{align}
To find the excitation point response function for illumination along the
$\mh{x}$ axis we need to change every $\mh{x}$ and $\mh{y}$ to $\mh{z}$ and
$-\mh{y}$, respectively. Rotating the spherical harmonics is not trivial,
but we show the relevant transformation matrix in Appendix \ref{sec:sh}. The
excitation point response function for illumination along the $\mh{x}$ axis is
given by
\begin{align}
  h^{\mh{x}, \text{ls}}_{\text{exc}}(\so{}; \mh{p}) \propto y_0^0(\so{}) + \frac{1}{\sqrt{5}}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{z})^2 - (\mh{p}\cdot\mh{y})^2]\left[\frac{\sqrt{3}}{2}y_2^0(\so{}) + \frac{1}{2}y_2^2(\so{})\right] + 2(\mh{p}\cdot\mh{z})(\mh{p}\cdot\mh{y})y_2^{-1}(\so{})\right\}.
\end{align}

To find the excitation transfer functions we need to calculate
\begin{align}
  H_{l, \text{exc}}^m = \int_{\mathbb{S}^2} d\so{}\, h_{\text{exc}}(\so{}) y_l^m(\so{}).
\end{align}
This calculation is straightforward now that we've expressed the excitation
point response functions in terms of spherical harmonics. The excitation
transfer functions are
\begin{align}
  H^{m,\mh{z}}_{l,\text{exc}}(\mh{p}) \propto\, &[2\bar{A} + (4/3)\bar{B}]\bar{D}\delta(l, m) - \frac{4\sqrt{5}}{15}\bar{B}\bar{D}\delta(l-2, m)(\so{}) + \nonumber\\ &\frac{4\bar{C}\bar{D}}{\sqrt{15}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]\delta(l-2, m-2) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})\delta(l-2, m+2)\right\},\\
  H^{m,\mh{z}, \text{ls}}_{l,\text{exc}}(\mh{p}) \propto\, &\delta(l,m) + \frac{1}{\sqrt{5}}\delta(l-2, m) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]\delta(l-2,m-2) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})\delta(l-2, m+2)\right\},\\
  H^{m,\mh{x}, \text{ls}}_{l, \text{exc}}(\mh{p}) \propto\, &\delta(l,m) + \frac{1}{\sqrt{5}}\delta(l-2, m) + \sqrt{\frac{3}{5}}\Bigg\{[(\mh{p}\cdot\mh{z})^2 - (\mh{p}\cdot\mh{y})^2]\left[\frac{\sqrt{3}}{2}\delta(l-2,m) + \frac{1}{2}\delta(l-2, m-2)\right]\nonumber\\ &\hspace{20em}+ 2(\mh{p}\cdot\mh{z})(\mh{p}\cdot\mh{y})\delta(l-2, m+1)\Bigg\}.     \end{align}

>>>>

We calculated several spatio-angular excitation point response functions in our
Optics Express paper \cite{chandler17} (in the paper we called this function the
excitation efficiency for a single dipole, but this is equivalent to the
excitation point spread function). If we place a spatially incoherent and
spatially uniform light source (or its image) in the aperture plane with the
optical axis along the $\mh{z}$ axis, then the excitation point response function
is given by
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\Theta, \Phi; \phi_{\text{exc}}) &= \bar{D}\{\bar{A} + \bar{B}\sin^{2}{\Theta} + \bar{C}\sin^{2}{\Theta} \cos{[2 (\Phi - \phi_{\text{exc}}})]\}\label{eq:scalarabs},
\end{align}
where
$\so{} = \cos\Phi\sin\Theta\mh{x} + \sin\Phi\sin\Theta\mh{y} + \cos\Theta\mh{z}$,
the illumination polarizer orientation is given by $\mh{p} = \cos\phi_{\text{exc}}\mh{x} + \sin\phi_{\text{exc}}\mh{y}$, and 
\begin{subequations}
\begin{align}
  \bar{A} &= \frac{1}{4} - \frac{3}{8} \cos{\alpha } + \frac{1}{8} \cos^{3}{\alpha },\\
  \bar{B} &= \frac{3}{16} \cos{\alpha } - \frac{3}{16} \cos^{3}{\alpha },\\
  \bar{C} &= \frac{7}{32} - \frac{3}{32} \cos{\alpha } - \frac{3}{32} \cos^{2}{\alpha } - \frac{1}{32} \cos^{3}{\alpha},\\
  \bar{D} &= \frac{4}{3(1 - \cos\alpha)},
\end{align}\label{eq:coefficients}%
\end{subequations}
where $\alpha = \arcsin(\text{NA}/n_o)$. I'm using bars on the constants to
avoid notation overlap with the previous note set. We can rewrite this
expression in terms of spherical harmonics as
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p}) \propto\, &[2\bar{A} + (4/3)\bar{B}]\bar{D}y_0^0(\so{}) - \frac{4\sqrt{5}}{15}\bar{B}\bar{D}y_2^0(\so{}) + \nonumber\\ &\frac{4\bar{C}\bar{D}}{\sqrt{15}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}. \label{eq:genpsf}
\end{align}
We can simplify this result further by redefining the constants and normalizing
\begin{align}
  h^{\mh{z}}_{\text{exc}}(\so{}; \mh{p}) &= y_0^0(\so{}) - \frac{1}{\sqrt{5}}\tilde{A}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\tilde{B}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}, \label{eq:genpsf}
\end{align}
where
\begin{subequations}
\begin{align}
  \tilde{A} &\equiv \cos^2(\alpha/2)\cos(\alpha),\\
  \tilde{B} &\equiv \frac{1}{12}(\cos^2\alpha + 4\cos\alpha + 7).
\end{align}\label{eq:coefficients}%
\end{subequations}

The light-sheet excitation point response function is given by Eq.
\ref{eq:genpsf} in the limit of small NA ($\tilde{A} \rightarrow 1$ and $\tilde{B} \rightarrow 1$) which gives
\begin{align}
  h^{\mh{z}, \text{ls}}_{\text{exc}}(\so{}; \mh{p}) \equiv \lim_{\alpha \rightarrow 0} h^{\mh{z}}_{\text{exc}}(\so{}) = y_0^0(\so{}) - \frac{1}{\sqrt{5}}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]y_2^2(\so{}) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})y_2^{-2}(\so{})\right\}.
\end{align}
To find the excitation point response function for illumination along the
$\mh{x}$ axis we need to change every $\mh{x}$ and $\mh{y}$ to $\mh{z}$ and
$-\mh{y}$, respectively. Rotating the spherical harmonics is not trivial,
but we show the relevant transformation matrix in Appendix \ref{sec:sh}. The
excitation point response function for illumination along the $\mh{x}$ axis is
given by
\begin{align}
  h^{\mh{x}, \text{ls}}_{\text{exc}}(\so{}; \mh{p}) = y_0^0(\so{}) - \frac{1}{\sqrt{5}}y_2^0(\so{}) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{z})^2 - (\mh{p}\cdot\mh{y})^2]\left[\frac{\sqrt{3}}{2}y_2^0(\so{}) + \frac{1}{2}y_2^2(\so{})\right] + 2(\mh{p}\cdot\mh{z})(\mh{p}\cdot\mh{y})y_2^{-1}(\so{})\right\}.
\end{align}

To find the excitation transfer functions we need to calculate
\begin{align}
  H_{l, \text{exc}}^m = \int_{\mathbb{S}^2} d\so{}\, h_{\text{exc}}(\so{}) y_l^m(\so{}).
\end{align}
This calculation is straightforward now that we've expressed the excitation
point response functions in terms of spherical harmonics. The excitation
transfer functions are
\begin{align}
  H^{m,\mh{z}}_{l,\text{exc}}(\mh{p}) &= \delta(l, m) - \frac{1}{\sqrt{5}}\tilde{A}\delta(l-2, m)(\so{}) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]\delta(l-2, m-2) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})\delta(l-2, m+2)\right\},\\
  H^{m,\mh{z}, \text{ls}}_{l,\text{exc}}(\mh{p}) &= \delta(l,m) - \frac{1}{\sqrt{5}}\delta(l-2, m) + \sqrt{\frac{3}{5}}\left\{[(\mh{p}\cdot\mh{x})^2 - (\mh{p}\cdot\mh{y})^2]\delta(l-2,m-2) - 2(\mh{p}\cdot\mh{x})(\mh{p}\cdot\mh{y})\delta(l-2, m+2)\right\},\\
  H^{m,\mh{x}, \text{ls}}_{l, \text{exc}}(\mh{p}) &= \delta(l,m) - \frac{1}{\sqrt{5}}\delta(l-2, m) + \sqrt{\frac{3}{5}}\Bigg\{[(\mh{p}\cdot\mh{z})^2 - (\mh{p}\cdot\mh{y})^2]\left[\frac{\sqrt{3}}{2}\delta(l-2,m) + \frac{1}{2}\delta(l-2, m-2)\right]\nonumber\\ &\hspace{20em}+ 2(\mh{p}\cdot\mh{z})(\mh{p}\cdot\mh{y})\delta(l-2, m+1)\Bigg\}.     \end{align}
